---
title: "R Notebook - Example for ploting raster data"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---



First, we import the library we will use in this example. The *raster* library will provide the functions to handle the raster object, while the *rasterVis* library will be used to plot the data. 

```{r}

library(raster)
library(rasterVis)
```

Next, we point to the working directory, list the files in the folder *Geodata*, and important each file individually. 

```{r}
setwd("/Volumes/GoogleDrive/My Drive/Teaching/2021/Data_Visualization/R/DataVisEss")

r_files <- list.files(path = "Geodata")


ndvi_1 <- raster(file.path("Geodata", r_files[2]))
names(ndvi_1) = "Wet_Season"
ndvi_2 <- raster(file.path("Geodata", r_files[3]))
names(ndvi_2) = "Dry_Season"

```

Let's take a look the files. They represent NDVI for southern Amazonia. 

```{r eval=TRUE}

levelplot(ndvi_1, margin=FALSE)
levelplot(ndvi_2, margin=FALSE)

```

Now, let's do something fun. *Stack* the two files and plot. 

```{r}
ndvi_1_2 <- stack(ndvi_1, ndvi_2)

levelplot(ndvi_1_2)

```

If you want to take the average of the two NDVI values, all you have to do is to take their *mean*. Done. 

```{r}
ndvi_mean <- mean(ndvi_1_2)
levelplot(ndvi_mean, margin = FALSE)

```

Wanna see the data distribution?

```{r}
hist(ndvi_1_2)
density(ndvi_1_2)
```
