---
title: "Week3_Raster_Precipitation"
output: html_document
editor_options: 
  chunk_output_type: console
---

We are going to use a couple libraries to import raster data (raster), handle file names (tidyverse), and plot nice figures (rasterVis).

```{r import_lib, echo=TRUE, warning=FALSE}

library(raster)
library(rasterVis)
library(tidyverse)


```

Here, we have to list the file path and names. I also listed only the file names, so I can subset it easily. 
```{r import_files, echo=TRUE}

#dir_files <- "/Volumes/GoogleDrive/My Drive/Teaching/2021/Data_Visualization/R/DataVisEss/Geodata/TRMM_Precip"
dir_files <- "~/Documents/Temp/Trmm"


path_files <- list.files(path = dir_files, 
                         full.names = TRUE,
                         pattern = "*nc")
files <- list.files(path = dir_files, 
                         full.names = FALSE,
                         pattern = "*.nc")
head(files)
```

This step is super important. It makes it very easy to choose which files to import. 

```{r organize_files, echo = TRUE}

table_names <- as_tibble(files) %>%
  bind_cols(path_files) %>%
  rename(File_path = `...2`) %>%
  mutate(value = str_sub(value, start = 1, end = 13)) %>%
  separate(value, c("Prod_name",  "yyyymmdd")) %>%
  select(yyyymmdd, File_path) %>%
  mutate(Yr = str_sub(yyyymmdd, start = 1, end = 4),
         Mo = str_sub(yyyymmdd, start = 5, end = 6),
         Da = str_sub(yyyymmdd, start = 7, end = 8)) %>%
  filter(Yr %in% 2015:2019)

table_names
```

Let's import the files and rename them. 

```{r import_rename, echo=TRUE}

trmm_stack <- stack(table_names$File_path)
extent(trmm_stack) = c(xmin = -124.102, xmax = -110.039, ymin = 34.453, ymax = 44.297)
crs(trmm_stack) <- CRS('+proj=longlat +ellps=WGS84')

trmm_stack
names(trmm_stack) <- table_names$yyyymmdd
trmm_stack

```

Taking the average of 5 yrs of data.

```{r basic_ex, echo = TRUE}

trmm_mean <- mean(trmm_stack)

#plot(trmm_mean)
levelplot(trmm_mean, margin = FALSE)

#cellStats(trmm_stack, "mean")
```

Let's plot the data in different ways. 

```{r colors, echo=TRUE}
library(RColorBrewer)
cutpts <- seq(0, 0.30, length = 10)

levelplot(trmm_mean, margin = FALSE, at=cutpts, cuts=11, pretty=T,
          col.regions=(rev(brewer.pal(10,"RdBu"))))

mapTheme <- rasterTheme(region=brewer.pal(8,"Blues"))
levelplot(trmm_mean, margin = FALSE, par.settings =mapTheme)
```

One option is to transform raster data into tabular data and do some stats. 

```{r from_raster_to_tab, echo = TRUE}

trmm_tab <- as.data.frame(trmm_stack) %>%
  as_tibble()

trmm_long <- trmm_tab %>%
  pivot_longer(names_to = "yyyymmdd", values_to = "values", cols = X20150101:X20191201) %>%
  mutate(Yr = str_sub(yyyymmdd, start = 2, end = 5),
         Mo = str_sub(yyyymmdd, start = 6, end = 7),
         Da = str_sub(yyyymmdd, start = 8, end = 9)) 
```


```{r ploting_data, echo=TRUE}

trmm_long %>%
  ggplot(aes(x = as.numeric(Mo), y = values)) +
  stat_summary(fun.data =  "mean_cl_boot", geom = "ribbon") +
  facet_wrap(~Yr) +
  theme_minimal()

trmm_long %>%
  ggplot(aes(x = Mo, y = values, color = as.factor(Yr))) +
  stat_summary(fun.data =  "mean_cl_boot") +
  theme_minimal()

trmm_long %>%
  ggplot(aes(x = as.numeric(Mo), 
             y = values, color = as.factor(Yr))) +
  geom_smooth() +
  theme_minimal()

```

```{r avg_by_groups}

trmm_2015 <-trmm_stack[[which(table_names$Yr == 2015)]]
avg_2015 <- mean(trmm_2015)

trmm_all <-trmm_stack[[which(table_names$Yr != 2015)]]
avg_all <- mean(trmm_all)

###
###
###
diff_15_all <- avg_2015 - avg_all

###
###
###

levelplot(diff_15_all, margin = FALSE)
levelplot(diff_15_all, margin = FALSE, contour=TRUE)
levelplot(diff_15_all, margin = TRUE, contour=TRUE)
levelplot(diff_15_all, margin = TRUE, contour=TRUE, par.settings = RdBuTheme)

```

```{r stackApply}
yr_trmm <- stackApply(trmm_stack, indices = as.integer(table_names$Yr), fun = mean, na.rm = TRUE)

levelplot(yr_trmm, margin = TRUE, contour=TRUE, par.settings = RdBuTheme)
levelplot(yr_trmm, margin = TRUE, contour=TRUE, par.settings = RdBuTheme) 
```

```{r drought_yr}
hist(diff_15_all)
```


```{r}

library(ggridges)

trmm_long %>%
  ggplot(aes(x = values, 
             y = Yr, 
             fill = stat(x))) +
  geom_density_ridges_gradient(alpha = 0.1) +
  scale_fill_viridis_c(name = "Prec (mm/hr)", option = "C") +
  #labs(title = 'Temperatures in Lincoln NE in 2016') +
  theme_ridges(font_size = 13, grid = TRUE) +
  theme(axis.title.y = element_blank())

trmm_long %>%
  ggplot(aes(x = values, 
             y = Yr,
             fill = as.factor(Yr))) +
  geom_density_ridges(alpha = 0.5) +
  theme_ridges(font_size = 13, grid = TRUE) +
  theme(axis.title.y = element_blank())

```
